---
title: "Camonsertib rBA - BP44676"
author: "Anh Nguyen Duc"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: journal
    highlight: pygments
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(PowerTOST) # This is to calculate the sample size for cross-over design with known coefficient of variation (CV)
require(MESS) # This is for function auc 
require(tidyverse)
require(knitr)
require(PK)
require(MASS)
require(table1)
require(haven)
require(readxl)
```

## Questions for Clinical Pharmacology Team
+ Should the AUC be compartmental or non-compartmental? $\Rightarrow$ Answer: non-compartmental.

+ Current data does not seem to have variable describing if dosing was under fed or fasted condition or if Prepare is planning to create such a variable? $\Rightarrow$ Answer: for TRESR Module 1 - monotherapy dose escalation, all data was under fasted condition until protocol v7. CV could be based on this. 

+ CV should only consider 160mg mono-dose, what about dosing schedule? $\Rightarrow$ Answer:all dosing schedule should be considered.

+ If mono-dose should be used $\rightarrow$ we should consider either or both of

  + TRESR Module 1: mono therapy dose escalation $\Rightarrow$ Answer:  all under fasted condition until protocol v7. CV could be based on this.
  
  + TRESR Module 2: mono therapy p2 (not sure how many patients we already have in the most recent Batch-28-20221108 transfer?)
  
  + Only mono-data should be used as confirmed by Marie-Laure.
  
  + Exact timing of dosing not available but for CV using nominal time would suffice.

## Read Data

### Data Derived by Roche PK Team
```{r}
nca_pk_dat <- read.csv(file = '../../0_Common/10_Data/4_PK_Parameters_Roche/RP3500_01_PKParameter.csv', header=T, sep = ';')

# see number of pts
nca_pk_dat$USUBJID %>% unique %>% length

# extract data with 160mg irrespective of dosing schedule
nca_pk_dat_160 <- nca_pk_dat %>% filter(TRTNAME %>% grepl(pattern='160', x=., fixed = TRUE))

# sanity check
nca_pk_dat_160$TRTNAME %>% unique

# see number of pts
nca_pk_dat_160$USUBJID %>% unique %>% length
```

### Raw SDTM Data from Repare
```{r}
# batch_sdtm <- '1_Batch-28-20221108'
batch_sdtm <- '3_Batch-29-20230109-Dec2022ClinicalData'

# For now we could only use SDTM

# lb <- read_sas(data_file = '../../10_Data/3_Data_Transferred/1_Batch-28-20221108/CO44446_TRESR_RP_3500_01/sdtm/lb.sas7bdat')
# 
# pf <- read_sas(data_file = '../../10_Data/3_Data_Transferred/1_Batch-28-20221108/CO44446_TRESR_RP_3500_01/sdtm/pf.sas7bdat')

ex <- read_sas(data_file = paste0('../../0_Common/10_Data/3_Data_Transferred/',batch_sdtm,'/CO44446_TRESR_RP_3500_01/sdtm/ex.sas7bdat') )

ec <- read_sas(data_file = paste0('../../0_Common/10_Data/3_Data_Transferred/',batch_sdtm,'/CO44446_TRESR_RP_3500_01/sdtm/ec.sas7bdat'))

pc <- read_sas(data_file = paste0('../../0_Common/10_Data/3_Data_Transferred/',batch_sdtm,'/CO44446_TRESR_RP_3500_01/sdtm/pc.sas7bdat'))

# meal data
ml <- read_sas(data_file = paste0('../../0_Common/10_Data/3_Data_Transferred/',batch_sdtm,'/CO44446_TRESR_RP_3500_01/sdtm/ml.sas7bdat'))

# map PC to EX to figure out which dose each of the pc lines corresponds to via first merging by timing of sample collection. This is needed to extract only mono160mg data

# remove empty concentration
pc  <- pc[pc$PCORRES != '',] 

# factorize the nominal PK collection timings for each dose
pc$PCTPT_factor <- pc$PCTPT %>% factor(levels=c('PRE-DOSE', '0.5 HOURS POST DOSE', '1 HOUR POST DOSE', '2 HOURS POST DOSE', '4 HOURS POST DOSE', '6 HOURS POST DOSE', '8 HOURS POST DOSE', '10 HOURS POST DOSE', '12 HOURS POST DOSE', '24 HOURS POST DOSE', '48 HOURS POST DOSE')) %>% ordered

tmp <- pc %>% group_by(USUBJID, VISIT) %>% summarise(PCRFTDTC_Date = PCRFTDTC[PCTPT_factor==PCTPT_factor%>%min])

pc$PCRFTDTC_Date <- sapply(1:nrow(pc), function(id) {
  sub <- pc$USUBJID[id]
  vis <- pc$VISIT[id]
  tmp$PCRFTDTC_Date[tmp$USUBJID==sub & tmp$VISIT==vis]
})

# Date conversion
pc$PCRFTDTC_Date <- pc$PCRFTDTC_Date %>% as.Date

# create nominal time
pc %>% dplyr::select(USUBJID, VISIT, PCTPT_factor, PCORRES, PCRFTDTC, PCRFTDTC_Date) %>% head(20) %>% kable

pc %>% dplyr::select(USUBJID, VISIT, PCTPT_factor, PCORRES, PCRFTDTC, PCRFTDTC_Date) %>% tail(20) %>% kable


ex$EXSTDTC_Date  <- ex$EXSTDTC %>% as.Date
ex$EXENDTC_Date  <- ex$EXENDTC %>% as.Date

ex %>% dplyr::select(USUBJID, EXSTDTC_Date, EXENDTC_Date, EXDOSE) %>% filter(USUBJID=='RP-3500-01-1002-0151')

tmp2 <- 
pc %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(EXDOSE = ex$EXDOSE[ex$USUBJID==unique(USUBJID) & (ex$EXSTDTC_Date<=unique(PCRFTDTC_Date) & ex$EXENDTC_Date>=unique(PCRFTDTC_Date) )]%>%unique %>% unlist) %>% na.omit

ex %>% dplyr::select(USUBJID, EXSTDTC_Date, EXENDTC_Date, EXDOSE) %>% filter(USUBJID=='RP-3500-01-1001-0003')

pc %>% dplyr::select(USUBJID, VISIT, PCTPT_factor, PCORRES, PCRFTDTC, PCRFTDTC_Date) %>% filter(USUBJID=='RP-3500-01-1001-0003')

pc$EXDOSE <- sapply(1:nrow(pc), function(id) {
  sub <- pc$USUBJID[id]
  ref <- pc$PCRFTDTC_Date[id]
  if (is.na(sub) | is.na(ref)) {
    re <- NA
  } else {
    re <- tmp2$EXDOSE[tmp2$USUBJID==sub & tmp2$PCRFTDTC_Date==ref] %>% unique  
  }
  if (length(re)==0) re <- NA
  re
}) %>% unlist

pc %>% dplyr::select(USUBJID, VISIT, PCTPT_factor, PCORRES, PCRFTDTC, PCRFTDTC_Date, EXDOSE) %>% head(20) %>% kable#filter(USUBJID=='RP-3500-01-1002-0151' & PCRFTDTC_Date=='2021-06-16')


## extract module 1 and 2 patients
pc <- pc %>% filter(USUBJID %>% substr(x=., start=12, stop=12) %in% c('1', '2'))
pc$USUBJID %>% substr(x=., start=12, stop=15) %>% unique

## extract patients with 160mg
pc160 <- pc %>% filter(EXDOSE==160 & !is.na(PCSTRESN))

## Compare pat id
pc160_id <- pc160$USUBJID %>% unique # 84
nca_pk_dat_160_id <- nca_pk_dat_160$USUBJID %>% unique # 77

# print out patients 
setdiff(pc160_id, nca_pk_dat_160_id) %>% kable
setdiff(nca_pk_dat_160_id, pc160_id) %>% kable

# so basically the PK team has fewer than 7 patients, which indeed had 160mg at some points
ex %>% dplyr::select(USUBJID, EXSTDTC_Date, EXENDTC_Date, EXDOSE) %>% filter(USUBJID %in% setdiff(pc160_id, nca_pk_dat_160_id), EXDOSE==160) %>% dplyr::select(USUBJID) %>% unique


# see food effect from pc
pc %>% filter(grepl("Fed", VISIT), PCRFTDTC!='', EXDOSE==160) %>% dplyr::select(USUBJID, VISIT, PCRFTDTC, EXDOSE) %>% as.data.frame %>% kable()

ml %>% as.data.frame %>% kable()

ml %>% filter(USUBJID %in%pc160_id) %>% as.data.frame %>% kable()

## so using VISIT column in pc or pc160 is a good idea
pc160_fasted <- pc160 %>% filter(!grepl("Fed", VISIT))
```

### Derived PK Data from Repare

Current data was transferred in batch 31

```{r}
nca_pk_dat_repare <- read_excel(path = '../../0_Common/10_Data/3_Data_Transferred/4_Batch-31-20230112/20230106 - Final Parameters Pivoted_120 patients_Roche (1).xlsx', sheet='Sheet1')
nca_pk_dat_repare$USUBJID <- nca_pk_dat_repare$USUBJID %>% sapply(FUN=function(x) {
  paste0('RP-3500-01-', substr(x = x, start=1, stop=4), '-',substr(x = x, start=5, stop=8))
})

nca_pk_dat_repare$AUC0_12 <- nca_pk_dat_repare$`AUC0-12`
nca_pk_dat_repare <- nca_pk_dat_repare %>% as.data.frame

# extract only module 1 and 2 patients
nca_pk_dat_repare <- nca_pk_dat_repare %>% filter(USUBJID %>% substr(x=., start=12, stop=12) %in% c('1', '2'))

# Remove data before dosing
# nca_pk_dat_repare <- nca_pk_dat_repare %>% filter(Day>0) # does not really matter
nca_pk_dat_repare_160 <- nca_pk_dat_repare %>% dplyr::filter(Dose =='160')

# Check if all pts in repare derivation is in Anh's derivation
nca_pk_dat_repare_160$USUBJID %>% unique %>% setdiff(x=., y=pc160_id) %>% kable
nca_pk_dat_repare_160$USUBJID %>% unique %>% setdiff(y=., x=pc160_id) %>% kable
nca_pk_dat_repare_160$USUBJID %>% unique %>% setdiff(x=., y=nca_pk_dat_160_id) %>% kable
nca_pk_dat_repare_160$USUBJID %>% unique %>% setdiff(y=., x=nca_pk_dat_160_id) %>% kable
```

## Trial Design

+ For most recent details see slide 8 or 11 of this deck: https://docs.google.com/presentation/d/1O0r1Mx-feQiC3BfKAdGQg2lojBAn-Fml/edit#slide=id.p2


+ Primary endpoints: $AUC_{\infty}$, $AUC_{\text{last}}$ and $C_{\max}$

## Estimation of Primary Endpoints for Each Patients and the Respective CV

### Summary of Relevant PK Parameters

#### Roche derivation

```{r}
table1(~AUCINF_obs%>%as.numeric+AUCINF_pred%>%as.numeric+AUCall%>%as.numeric+AUClast%>%as.numeric+Cmax%>%as.numeric+Clast%>%as.numeric, data=nca_pk_dat_160)
```

#### Repare derivation

```{r}
table1(~AUCINF_obs%>%as.numeric+AUC0_12%>%as.numeric+AUClast%>%as.numeric+Cmax%>%as.numeric, data=nca_pk_dat_repare_160)
```

### Estimate $AUC_{\infty}$

**Questions**

+ Should we use `AUCINF_pred` or `AUCINF_obs`? In any case they are virtually similar as the plot below.

```{r}
with(nca_pk_dat_160, plot(x = AUCINF_obs %>% as.numeric, y = AUCINF_pred %>% as.numeric, col='blue'))
abline(a=0, b=1)
```

**Estimation**
<!-- To-do 
+ For self derivation, use function nca in package PK
-->

```{r}
aucinf_anh <- pc160 %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(AUCINF_obs=nca.complete(conc=PCSTRESN,                                     time=((PCDTC%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))-
       (PCDTC[1]%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))) %>% as.numeric/3600
  )$est['AUC to infinity',1]
)

aucinf_anh <- aucinf_anh %>% filter(AUCINF_obs >= 0)

aucinf_anh_fasted <- pc160_fasted %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(AUCINF_obs=nca.complete(conc=PCSTRESN,                                     time=((PCDTC%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))-
       (PCDTC[1]%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))) %>% as.numeric/3600
  )$est['AUC to infinity',1]
)

aucinf_anh_fasted <- aucinf_anh_fasted %>% filter(AUCINF_obs >= 0)
```

**Check if data is normally or log-normally distributed**

```{r}
# For Roche PK team derived data
with(nca_pk_dat_160, {
  x <- AUCINF_obs %>% as.numeric %>% na.omit
  hist(AUCINF_obs %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_160$AUCINF_obs_log <- nca_pk_dat_160$AUCINF_obs %>% as.numeric %>% log

with(nca_pk_dat_160, {
  x <- AUCINF_pred %>% as.numeric %>% na.omit
  hist(AUCINF_pred %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_160$AUCINF_pred_log <- nca_pk_dat_160$AUCINF_pred %>% as.numeric %>% log

# For self-derived data
with(aucinf_anh, {
  x <- AUCINF_obs %>% as.numeric %>% na.omit
  hist(AUCINF_obs %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)
aucinf_anh$AUCINF_obs_log <- aucinf_anh$AUCINF_obs %>% log

with(aucinf_anh_fasted, {
  x <- AUCINF_obs %>% as.numeric %>% na.omit
  hist(AUCINF_obs %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)
aucinf_anh_fasted$AUCINF_obs_log <- aucinf_anh_fasted$AUCINF_obs %>% log

# For repare-derived data 
with(nca_pk_dat_repare_160, {
  x <- AUCINF_obs %>% as.numeric %>% na.omit
  hist(AUCINF_obs %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_repare_160$AUCINF_obs_log <- nca_pk_dat_repare_160$AUCINF_obs %>% as.numeric %>% log
```

**Estimate CV**

<!-- To-do
+ Use anova to first get the mse = msq / df for residuals and get CV using mse2CV function in package PowerTOST
-->

```{r}
# for Roche PK team derived data
tmp_anova <- aov(AUCINF_obs_log ~ USUBJID, data = nca_pk_dat_160)
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_aucinf_obs_log <- mse2CV(mse))

# repeat for AUCIINF_pred
tmp_anova <- aov(AUCINF_pred_log ~ USUBJID, data = nca_pk_dat_160)
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_aucinf_pred_log <- mse2CV(mse))

# repeat for self-derived data
tmp_anova2 <- aov(AUCINF_obs_log ~ USUBJID, data = aucinf_anh)
mse2       <- (tmp_anova2$residuals ^ 2 %>% sum) / tmp_anova2$df.residual

(cv_aucinf_obs_log2 <- mse2CV(mse2))

tmp_anova2_fasted <- aov(AUCINF_obs_log ~ USUBJID, data = aucinf_anh_fasted)
mse2_fasted       <- (tmp_anova2_fasted$residuals ^ 2 %>% sum) / tmp_anova2_fasted$df.residual

(cv_aucinf_obs_log2_fasted <- mse2CV(mse2_fasted))

# For self-derived data and consider only pts appearing in repare's pk sets
tmp_anova3 <- aov(AUCINF_obs_log ~ USUBJID, data = aucinf_anh %>% filter(USUBJID %in% (nca_pk_dat_repare_160$USUBJID %>% unique)) )
mse3       <- (tmp_anova3$residuals ^ 2 %>% sum) / tmp_anova3$df.residual

(cv_aucinf_obs_log3 <- mse2CV(mse3))

tmp_anova3_fasted <- aov(AUCINF_obs_log ~ USUBJID, data = aucinf_anh_fasted %>% filter(USUBJID %in% (nca_pk_dat_repare_160$USUBJID %>% unique)))
mse3_fasted       <- (tmp_anova3_fasted$residuals ^ 2 %>% sum) / tmp_anova3_fasted$df.residual

(cv_aucinf_obs_log3_fasted <- mse2CV(mse3_fasted))

# repeat for repare-derived data
tmp_anova <- aov(AUCINF_obs_log ~ USUBJID, data = nca_pk_dat_repare_160)
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_aucinf_obs_log_repare <- mse2CV(mse))
```

### Estimate $AUC_{\text{last}}$

**Questions**

+ Should we use `AUClast` or `AUCall`? In any case they are virtually similar as the plot below.

```{r}
with(nca_pk_dat_160, plot(x = AUClast %>% as.numeric, y = AUCall %>% as.numeric, col='blue'))
abline(a=0, b=1)
```

**Estimation**
<!-- To-do 
+ For self derivation, use function nca in package PK
-->

```{r}
auclast_anh <- pc160 %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(AUClast=nca.complete(conc=PCSTRESN,                                     time=((PCDTC%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))-
       (PCDTC[1]%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))) %>% as.numeric/3600
  )$est['AUC to tlast',1]
)


auclast_anh_fasted <- pc160_fasted %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(AUClast=nca.complete(conc=PCSTRESN,                                     time=((PCDTC%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))-
       (PCDTC[1]%>% as.POSIXct(format="%Y-%m-%dT%H:%M"))) %>% as.numeric/3600
  )$est['AUC to tlast',1]
)
```

**Check if data is normally or log-normally distributed**

```{r}
with(nca_pk_dat_160, {
  x <- AUClast %>% as.numeric %>% na.omit
  hist(AUClast %>% as.numeric %>% na.omit, freq=F, ylim=c(0, 2e-5))
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_160$AUClast_log <- nca_pk_dat_160$AUClast %>% as.numeric %>% log

with(nca_pk_dat_160, {
  x <- AUCall %>% as.numeric %>% na.omit
  hist(AUCall %>% as.numeric %>% na.omit, freq=F, ylim=c(0, 2e-5))
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_160$AUCall_log <- nca_pk_dat_160$AUCall %>% as.numeric %>% log

# self derived data
with(auclast_anh , {
  x <- AUClast %>% as.numeric %>% na.omit
  hist(AUClast %>% as.numeric %>% na.omit, freq=F, ylim=c(0, 2e-5))
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

auclast_anh$AUClast_log <- auclast_anh$AUClast %>% log


with(auclast_anh_fasted , {
  x <- AUClast %>% as.numeric %>% na.omit
  hist(AUClast %>% as.numeric %>% na.omit, freq=F, ylim=c(0, 2e-5))
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

auclast_anh_fasted$AUClast_log <- auclast_anh_fasted$AUClast %>% log
```

**Estimate CV**

<!-- To-do
+ Use anova to first get the mse = msq / df for residuals and get CV using mse2CV function in package PowerTOST
-->

```{r}
tmp_anova <- aov(AUClast_log ~ USUBJID, data = nca_pk_dat_160)
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_auclast_log <- mse2CV(mse))

# repeat for AUCIINF_pred
tmp_anova <- aov(AUCall_log ~ USUBJID, data = nca_pk_dat_160)
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_aucall_log <- mse2CV(mse))

# repeat for self-derived AUClast
tmp_anova2 <- aov(AUClast_log ~ USUBJID, data = auclast_anh)
mse2       <- (tmp_anova2$residuals ^ 2 %>% sum) / tmp_anova2$df.residual

(cv_auclast_log2 <- mse2CV(mse2))

tmp_anova2_fasted <- aov(AUClast_log ~ USUBJID, data = auclast_anh_fasted)
mse2_fasted       <- (tmp_anova2_fasted$residuals ^ 2 %>% sum) / tmp_anova2_fasted$df.residual

(cv_auclast_log2_fasted <- mse2CV(mse2_fasted))
```

### Estimate $C_{\max}$

**Estimation**

<!-- To-do 
+ For self derivation, use function nca in package PK
-->
```{r}
cmax_anh <- pc160 %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(Cmax=PCSTRESN[PCTPT!='PRE-DOSE'] %>% max(na.rm = T), VISIT=VISIT[PCTPT!='PRE-DOSE'] %>%unique)

cmax_anh$Cycle <- sapply(cmax_anh$VISIT, FUN = function(x) {
  case_when(
    x %>% grepl(x=., pattern='Fed Day') ~ 1,
    x %>% grepl(x=., pattern='Cycle 1') ~ 1,
    x %>% grepl(x=., pattern='Cycle 2') ~ 2,
    x %>% grepl(x=., pattern='Cycle 3') ~ 3,
    x %>% grepl(x=., pattern='Cycle 4') ~ 4
  )  
})

cmax_anh$Day <- sapply(cmax_anh$VISIT, FUN = function(x) {
  case_when(
    x %>% grepl(x=., pattern='Fed Day') ~ -3,
    x %>% grepl(x=., pattern='Day 1') ~ 1,
    x %>% grepl(x=., pattern='Day 2') ~ 2,
    x %>% grepl(x=., pattern='Day 3') ~ 3,
    x %>% grepl(x=., pattern='Day 4') ~ 4,
    x %>% grepl(x=., pattern='Day 5') ~ 5
  )  
})

cmax_anh_fasted <- pc160_fasted %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(Cmax=PCSTRESN[PCTPT!='PRE-DOSE'] %>% max(na.rm = T), VISIT=VISIT[PCTPT!='PRE-DOSE'] %>%unique)

cmax_anh_fasted$Cycle <- sapply(cmax_anh_fasted$VISIT, FUN = function(x) {
  case_when(
    x %>% grepl(x=., pattern='Fed Day') ~ 1,
    x %>% grepl(x=., pattern='Cycle 1') ~ 1,
    x %>% grepl(x=., pattern='Cycle 2') ~ 2,
    x %>% grepl(x=., pattern='Cycle 3') ~ 3,
    x %>% grepl(x=., pattern='Cycle 4') ~ 4
  )  
})

cmax_anh_fasted$Day <- sapply(cmax_anh_fasted$VISIT, FUN = function(x) {
  case_when(
    x %>% grepl(x=., pattern='Fed Day') ~ -3,
    x %>% grepl(x=., pattern='Day 1') ~ 1,
    x %>% grepl(x=., pattern='Day 2') ~ 2,
    x %>% grepl(x=., pattern='Day 3') ~ 3,
    x %>% grepl(x=., pattern='Day 4') ~ 4,
    x %>% grepl(x=., pattern='Day 5') ~ 5
  )  
})
```

**Check if data is normally or log-normally distributed**

```{r}
# Roche PK team derived data
with(nca_pk_dat_160, {
  x <- Cmax %>% as.numeric %>% na.omit
  hist(Cmax %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_160$Cmax_log <- nca_pk_dat_160$Cmax %>% as.numeric %>% log

# Anh's derivation
with(cmax_anh, {
  x <- Cmax %>% as.numeric %>% na.omit
  hist(Cmax %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

cmax_anh$Cmax_log <- cmax_anh$Cmax %>% as.numeric %>% log


with(cmax_anh_fasted, {
  x <- Cmax %>% as.numeric %>% na.omit
  hist(Cmax %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

cmax_anh_fasted$Cmax_log <- cmax_anh_fasted$Cmax %>% as.numeric %>% log


# Repare derived data
with(nca_pk_dat_repare_160, {
  x <- Cmax %>% as.numeric %>% na.omit
  hist(Cmax %>% as.numeric %>% na.omit, freq=F)
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_repare_160$Cmax_log <- nca_pk_dat_repare_160$Cmax %>% as.numeric %>% log

# compare Anh's drivation vs Repare's 
tmp <- cmax_anh #%>% filter(USUBJID %in% (nca_pk_dat_repare_160$USUBJID %>% unique))

tmp_join <-
inner_join(x=tmp %>% dplyr::select(USUBJID, PCRFTDTC_Date, Cmax, Cycle, Day), y=nca_pk_dat_repare_160 %>% dplyr::select(USUBJID, Cycle, Day, Cmax), by = c('USUBJID', 'Cycle', 'Day')) 

tmp_join %>% kable
```  

**Estimate CV**

<!-- To-do
+ Use anova to first get the mse = msq / df for residuals and get CV using mse2CV function in package PowerTOST
-->

```{r}
# Roche PK team derived data
tmp_anova <- aov(Cmax_log ~ USUBJID, data = nca_pk_dat_160)
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_cmax_log <- mse2CV(mse))

# Anh team derived data
tmp_anova2 <- aov(Cmax_log ~ USUBJID, data = cmax_anh)
mse2       <- (tmp_anova2$residuals ^ 2 %>% sum) / tmp_anova2$df.residual

(cv_cmax_log2 <- mse2CV(mse2))

tmp_anova2_fasted <- aov(Cmax_log ~ USUBJID, data = cmax_anh_fasted)
mse2_fasted       <- (tmp_anova2_fasted$residuals ^ 2 %>% sum) / tmp_anova2_fasted$df.residual

(cv_cmax_log2_fasted <- mse2CV(mse2_fasted))

# For self-derived data and consider only pts appearing in repare's pk sets
tmp_anova3 <- aov(Cmax_log ~ USUBJID, data = cmax_anh %>% filter(USUBJID %in% (nca_pk_dat_repare_160$USUBJID %>% unique)) )
mse3       <- (tmp_anova3$residuals ^ 2 %>% sum) / tmp_anova3$df.residual

(cv_cmax_log3 <- mse2CV(mse3))

tmp_anova3_fasted <- aov(Cmax_log ~ USUBJID, data = cmax_anh_fasted %>% filter(USUBJID %in% (nca_pk_dat_repare_160$USUBJID %>% unique)) )
mse3_fasted       <- (tmp_anova3_fasted$residuals ^ 2 %>% sum) / tmp_anova3_fasted$df.residual

(cv_cmax_log3_fasted <- mse2CV(mse3_fasted))

# Repare derived data
tmp_anova <- aov(Cmax_log ~ USUBJID, data = nca_pk_dat_repare_160 #%>%filter(USUBJID!='RP-3500-01-1001-0007')
                 )
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_cmax_log_repare <- mse2CV(mse))

# using the joint data set
tmp_anova4 <- aov(I(Cmax.x %>% log) ~ USUBJID, data = tmp_join)
mse4       <- (tmp_anova4$residuals ^ 2 %>% sum) / tmp_anova4$df.residual

(cv_cmax_log4 <- mse2CV(mse4))

tmp_anova5 <- aov(I((Cmax.y %>% as.numeric)%>%log) ~ USUBJID, data = tmp_join %>% filter(!USUBJID %in% c('RP-3500-01-3003-0217', 'RP-3500-01-1001-0007')))
mse5       <- (tmp_anova5$residuals ^ 2 %>% sum) / tmp_anova5$df.residual

(cv_cmax_log5 <- mse2CV(mse5))
```


### Estimate $C_{\text{last}}$

**Estimation**

<!-- To-do 
+ For self derivation, use function nca in package PK
-->
```{r}
cvlast_anh <- pc160 %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(Clast=PCSTRESN[PCTPT!='PRE-DOSE'] %>% na.omit %>% tail(1))

cvlast_anh_fasted <- pc160_fasted %>% group_by(USUBJID, PCRFTDTC_Date) %>% summarise(Clast=PCSTRESN[PCTPT!='PRE-DOSE'] %>% na.omit %>% tail(1))
```

**Check if data is normally or log-normally distributed**

```{r}
with(nca_pk_dat_160, {
  x <- Clast %>% as.numeric %>% na.omit
  hist(Clast %>% as.numeric %>% na.omit, freq=F, ylim=c(0,20e-4))
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:max(x),fit[1],fit[2]), lwd=3)
  }
)

nca_pk_dat_160$Clast_log <- nca_pk_dat_160$Clast %>% as.numeric %>% log

# Anh's derivation
with(cvlast_anh, {
  x <- Clast %>% as.numeric %>% na.omit
  hist(Clast %>% as.numeric %>% na.omit, freq=F, ylim=c(0,20e-4))
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:last(x),fit[1],fit[2]), lwd=3)
  }
)

cvlast_anh$Clast_log <- cvlast_anh$Clast %>% as.numeric %>% log


with(cvlast_anh_fasted, {
  x <- Clast %>% as.numeric %>% na.omit
  hist(Clast %>% as.numeric %>% na.omit, freq=F, ylim=c(0,20e-4))
  fit<-fitdistr(x,"log-normal")$estimate
  lines(dlnorm(0:last(x),fit[1],fit[2]), lwd=3)
  }
)

cvlast_anh_fasted$Clast_log <- cvlast_anh_fasted$Clast %>% as.numeric %>% log
```  

**Estimate CV**

<!-- To-do
+ Use anova to first get the mse = msq / df for residuals and get CV using mse2CV function in package PowerTOST
-->

```{r}
tmp_anova <- aov(Clast_log ~ USUBJID, data = nca_pk_dat_160)
mse       <- (tmp_anova$residuals ^ 2 %>% sum) / tmp_anova$df.residual

(cv_clast_log <- mse2CV(mse))

tmp_anova2 <- aov(Clast_log ~ USUBJID, data = cvlast_anh)
mse2       <- (tmp_anova2$residuals ^ 2 %>% sum) / tmp_anova2$df.residual

(cv_clast_log2 <- mse2CV(mse2))

tmp_anova2_fasted <- aov(Clast_log ~ USUBJID, data = cvlast_anh_fasted)
mse2_fasted       <- (tmp_anova2_fasted$residuals ^ 2 %>% sum) / tmp_anova2_fasted$df.residual

(cv_clast_log2_fasted <- mse2CV(mse2_fasted))
```

## Sample Size Estimation

### Requirements

```{r }
target_power <- .8 #.85, .9
theta0      <- .95 #.9
alpha       <- .1

log_scale <- T

CVs <- c('aucinf_obs_log' = cv_aucinf_obs_log,
         'aucinf_pred_log' = cv_aucinf_pred_log,
         'aucinf_obs_log2' = cv_aucinf_obs_log2,
         'aucinf_obs_log2_fasted' = cv_aucinf_obs_log2_fasted,
         'aucinf_obs_log3' = cv_aucinf_obs_log3,
         'aucinf_obs_log3_fasted' = cv_aucinf_obs_log3_fasted,         
         'aucinf_obs_log_repare'  = cv_aucinf_obs_log_repare,
         'aucall_log' = cv_aucall_log,
         'auclast_log' = cv_auclast_log,
         'auclast_log2' = cv_auclast_log2,
         'auclast_log2_fasted' = cv_auclast_log2_fasted,
         'cmax_log' = cv_cmax_log,
         'cmax_log2' = cv_cmax_log2,
         'cmax_log2_fasted' = cv_cmax_log2_fasted,
         'cmax_log3' = cv_cmax_log3,
         'cmax_log3_fasted' = cv_cmax_log3_fasted,         
         'cmax_log_repare' = cv_cmax_log_repare,
         'cmax_log5'=cv_cmax_log5,
         'clast_log' = cv_clast_log,
         'nominal' = .27)

N_total_2x2 <- sapply(CVs, function(x) {
    sampleN.TOST(CV = x, alpha = alpha, targetpower = target_power, theta0=theta0, logscale = log_scale, design = '2x2', method = 'exact', robust = T, print = F)$`Sample size`
}) # "robust" is not relevant for 2x2 design

N_total_3x3 <- sapply(CVs, function(x) {
    sampleN.TOST(CV = x, alpha = alpha, targetpower = target_power, theta0=theta0, logscale = log_scale, design = '3x3', method = 'exact', robust = T, print = F)$`Sample size`
}) # for 3x3 "robust" to T tends to give  slightly larger sample size

# Print results
tmp_tab <-
rbind(
  CVs,
  'N_total_2x2' = N_total_2x2,
  'N_total_3x3' = N_total_3x3
) %>% t

rownames(tmp_tab) <- CVs %>% names
tmp_tab %>% kable()
```

## Power for 24 Subjects and Various BE Margins and CV of 0.27
```{r }
theta1 <- c(.8, .775, .75, .725, .7)
theta2 <- 1/theta1
maxCV  <- .27 # as written in protocol #CVs[CVs<=.5] %>% max
n <- 24
pows <- 
sapply(theta1, function(x) power.TOST(CV = maxCV, alpha = alpha, n=n, theta0=theta0, theta1=x, logscale = log_scale, design = '3x3', method = 'exact', robust = T))

cbind('low'=theta1, 'upp'=theta2, 'pow'=pows) %>% kable(digits = 3, row.names = F, caption = paste0('Power for various BE interval for ', n, ' subjects and', ' CV=', maxCV))
``` 

<!-- ## Code-testing Chunks to be Commented Out -->

<!-- ```{r } -->
<!-- # log_scale <- T -->
<!-- # mean <- c(0.1262, 0.1406, 0.1655,  0.1392, 0.1508 ) -->
<!-- # sd   <- c(0.0634, 0.0881, 0.05658, 0.0600, 0.07123) -->
<!-- #  -->
<!-- # CV   <- sd/mean -->
<!-- # if (log_scale) CV <- sqrt(exp(sd^2)-1) -->

<!-- log_scale <- T -->
<!-- CV <- c(.05, .1, .15, .2, .25, .3, .35, .4, .45, .5) -->
<!-- ``` -->

<!-- ### Results for AB x BA - no food effect -->
<!-- ```{r } -->
<!-- N_total_2x2 <- sapply(CV, function(x) { -->
<!--     sampleN.TOST(CV = x, alpha = .1, targetpower = .8, logscale = log_scale, design = '2x2', method = 'exact', robust = T, print = F)$`Sample size` -->
<!-- }) # "robust" is not relevant for 2x2 design -->

<!-- ``` -->


<!-- ### Results for T_fa T_fe C_fa x C_fa T_fa T_fe x T_fe C_fa T_fa -->
<!-- ```{r } -->
<!-- N_total_3x3 <- sapply(CV, function(x) { -->
<!--     sampleN.TOST(CV = x, alpha = .1, targetpower = .8, logscale = log_scale, design = '3x3', method = 'exact', robust = T, print = F)$`Sample size` -->
<!-- }) # for 3x3 "robust" to T tends to give  slightly larger sample size -->
<!-- ``` -->

<!-- ### Display -->
<!-- ```{r } -->
<!-- rbind( -->
<!--   'CV_logscale' = CV, -->
<!--   'N_total_2x2' = N_total_2x2, -->
<!--   'N_total_3x3' = N_total_3x3 -->
<!-- ) %>% kable() -->
<!-- ``` -->

## Session Info

```{r}
sessionInfo()
```
